---
title: "gprofiler2.addon : quick guide"
output: rmarkdown::html_vignette
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "Gwenaëlle Lemoine"
vignette: >
  %\VignetteIndexEntry{quick-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*gprofiler2.addon is a small package to add functionnalities to [gprofiler2](https://github.com/cran/gprofiler2)*

__________________

Suppose we want to use a gmt file which isn't available into `gprofiler2::gost`, for exemple those from [MSigDB](http://software.broadinstitute.org/gsea/msigdb/collections.jsp).
The `h.all.v6.2.entrez.gmt` is available into the addon :

```{r gmt_detail, message = FALSE}
library(gprofiler2.addon)
path_to_gmt <- system.file("extdata", "h.all.v6.2.entrez.gmt", package = "gprofiler2.addon", mustWork = TRUE)
gmt <- gprofiler2.addon::format_gmt(path_to_gmt)
```
**GMT content example**
`r knitr::kable(data.frame(Term = names(gmt[c(6,11)]), IDs = paste0(gmt[c(6,11)], sep=", ")))`

## Using gost with custom gmt
```{r gost_custom}
query <- c("26118", "5837", "6781", "23036", "694", "123", "1466", "7436",
          "23210", "2131", "2152", "5165", "55139", "7360", "229", "8614",
          "54206", "2027", "10957", "3162", "5228", "26330", "9435", "55076")
gost_custom_gmt_res <- gost_custom_gmt(query, custom_gmt = path_to_gmt)
```
**Result table** (subset of columns)
`r knitr::kable(gost_custom_gmt_res$result[,c("p_value", "term_id", "source", "term_name", "parents")])`

## Aggregating results 
One could also want to merge results from custom gmt with classic gost
```{r gost_aggreg, message = FALSE}
gost_classic_res <- gprofiler2::gost(query, numeric_ns = "ENTREZGENE_ACC")
aggreg_res <- gost_aggreg_res(gost_classic_res, gost_custom_gmt_res)
```
**Result table** (subset of columns)
`r knitr::kable(aggreg_res$result[,c("p_value", "term_id", "source", "term_name", "parents")])`

## Plotting together
Aggregated results can be plot together with `gprofiler2::gostplot` since `gprofiler2::gost_aggreg_res` output respect the same format.
```{r gostplot, fig.width=7, fig.height=5}
gprofiler2::gostplot(aggreg_res)
```

